Visually pleasing knot projections

### 2022-06-15 inst/jma_rebuttal_review3.tex

\textcolor{blue}{Canonical URL now added\\ \\}

Reviewer \#2: Firstly I would like to apologise for my late response on
this. Reviewing this edited paper had not got onto my radar.

Secondly all the points in my original report have now been
addressed. I do feel that the literature review could have gone a
little bit further beyond my stated suggestions, but it is now
sufficient to give the paper more context.

\textcolor{blue}{This revision incorporates many referee's suggestions
  into the literature review.  I would agree that the literature
  review is tightly focussed on the aesthetics of knot diagrams.
  Although it is clear that many workers place high value on pleasing
  diagrams, few sources explicitly address the issue.  I would of
  course be happy to cite any suggested references.}

### 2022-06-15 inst/jma_rebuttal_review3.tex

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

(Below, the reviewer's comments are in black, and my replies to the
issues are in \textcolor{blue}{blue}.  I have indicated changes to the
manuscript where appropriate.

\section*{Rebuttal to second third review: overall response}

\textcolor{blue}{
Short story: the reviewers make reasonable and constructive comments.
I have accommodated the comments with  with rewording.  The resulting
document is, I believe, stronger and more scholarly than before and I
recommend it to you.}

@@

Reviewer \#1: The paper has improved in terms of its overall
presentation in the context of knot renderings, and many detailed
fixes have been made as suggested by the reviewers.

However, the paper still has the major weakness of being too much a
somewhat ad-hoc technical manual on ``How-to-use-R and Inkplot to get
some decent knot plots" by using the software package [27].

The paper is lacking good discussions of the key issues and
explanations of the choices made, specifically: What makes a knot plot
attractive, followed by a thorough discussion of how to best set up a
knot representation and an optimization procedure to achieve the
desired qualities?

\textcolor{blue}{I appreciate the sentiment, which I take to be a
  comment on the large amount of programming-oriented detail in the
  submission.  This reviewer evidently regards the knots in the
  submission as ``decent'', which I take to be a great compliment.  It
  is often difficult to know the correct balance between an overly
  technical description [which indeed often overlaps with the aims of
    a technical manual] and conceptual big-picture thinking.  However,
  I note that many of the comments below are essentially requests for
  clarifications of minor technical details of the implementation: and
  this would indicate that perhaps the balance should be shifted back
  toward the technical manual end of the spectrum.  The other
  observation I would make here is that the software package [27] is
  intended to be part of the submission (it is released under the GPL
  and is thus available for public use) and with this understanding,
  even a downbeat technical manual would have scholarly value in
  conjunction with the new software it describes.  I want the editor
  and reviewer to know that in addressing the concerns below I have
  tried hard to answer not just ``what'' but also the ``why'' of the
  implementation.  \\ \\ Nevertheless, the cue above is a good one.
  The two questions: (a) {\em what makes a knot plot attractive?}, and
  (b) {\em how best ... to achieve this?}  are indeed profound, and
  were not addressed directly in the old manuscript.  In the revision
  I have prominently stated the questions in the introduction (section
  1.2), and given a provocative, if brief, outline of my view.\\ \\}

@@

The problem starts with the basic curve representation:

Why use Bezier curves?

\textcolor{blue}{Most of the answer is in the manuscript: ``A Bezier
  curve is a visually pleasing polynomial path that can be used to
  specify the path of a knot projection; here cubic Bezier curves are
  used''.  However, as observed, there are alternatives.  I have added
  the following sentence: ``Bezier curves are a natural choice for
  working with mathematical knots for several reasons: they are
  familiar to many graphical workers; they have a natural and
  intuitive control system (usually called ``handles''), and are
  implemented in many graphical software suites''.  I do not wish to
  suggest that Bezier curves are the only parametrizations for working
  with knots; but they are a natural starting point, and AFAICS they
  have no serious disadvantages in this context, nor is there any
  obvious superior method.\\ \\}

@@

And, if forced by the chosen tool to use Bezier curves, why not place
nodes at the crossing points, (and one or two more for large loopy
lobes between subsequent crossing points)?  -- This would allow the
user to place the crossing points far enough apart to start with, and
this would make it easy to set the Bezier handles to force right-angle
crossings.

\textcolor{blue}{Most of this issue is discussed below, at the comment
  for page 4, lines 53-54.  But here I would observe that it is not at
  all easy to manipulate Bezier handles on two distinct nodes [which
    happen to coincide spatially] to result in their two respective
  lines being orthogonal.  Observe that the overstrand and understrand
  of a crossing point may have (for good reasons!) very different
  curvatures.\\ \\}

Also, the implementation in the ``badness function" seems far from
optimal.  ``Total crossing angles()" does not address a key problem
directly: If one wants to avoid highly acute crossing angles, then
this can be addressed in a much more sensitive way with a function
such as 1/ sqr( cross.angle ).  This would go to infinity in the worst
case!  It also corresponds more directly to the way that excessive
curvature is avoided by minimizing ``bending energy," which is the
integral over the square of curvature -- or: 1/ sqr( curve-radius).
These are some ``math" issues that readers of JMA would be interested
in.

\textcolor{blue}{Function {\tt total\_crossing\_angles()} was poorly
  named.  With crossing angles $\theta_i$, $i=1,2,\ldots k$ the
  function actually returns $\sum_{i=1}^k\left(\cos\theta_i\right)^2$.
  \\ \\ The function is now named {\tt
    total\_crossing\_angle\_badness()}, which more accurately
  describes its purpose.\\ \\}

@@

Some more detailed comments:

Page 3 line 42: ``...knots possess a line of symmetry (at least, the
diagrams do if the breaks are ignored), which ..."  -$>>$ ``...knots
possess a C2 rotation axis, which ..."

\textcolor{blue}{I do not believe that describing the symmetry in
  question as a C2 rotation axis is appropriate.  A C2 axis refers to
  a three-dimensional object which is unchanged by a $180^\circ$
  rotation about a particular axis in 3D space.  I am considering the
  knot path here, which is a plane figure.  Referring to a C2 axis
  might suggest to readers that I am considering a 3D figure [e.g. the
    result of taking a 2D figure and perturbing strands slightly at a
    crossing point to avoid self-intersection].  But this particular 3D
  figure might not possess a C2 axis, a potentially confusing (and
  anyway irrelevant) fact.  On balance, I believe that mirror symmetry
  is the appropriate concept here, but the original sentence was
  poorly worded.\\ \\ OLD TEXT: ...possess a line of symmetry (at
  least if...ignored)\\ \\ NEW TEXT: ...possess mirror symmetry (at
  least if...ignored)\\ \\ }

@@

Page 4 lines 22-25: ``One plausible technique for creating knot
projections is to consider a two-dimensional projection of a knot's
embedding E in R3. However, this approach often results in poor visual
appearance: cusps or other displeasing effects can occur."  -$>>$ True!
- But what is the alternative, if the user has an entangled physical
knot in her hands?

\textcolor{blue}{Being able to render a nice optimized diagram
  corresponding to a physical knot that one has in one's hands is a
  reasonable wish.  But not one that I can fulfil.  Indeed, the
  package is predicated upon the notion that the diagram is the
  ``real'' object, and the (3D) knot itself is to some extent
  irrelevant.  Thus much of the package {\em prevents} the
  optimization routine from making Reidermeister moves [that is,
    changes that affect the topology of the diagram but not that of
    the 3D knot].  The operational reason for this is that the user
  specifies a particular diagram which the software then optimizes for
  appearance.  The manuscript includes two diagrams of $10_{125}$, the
  Perko pair, which show that sometimes one wants to preserve the
  topology of the diagram.\\ \\}

@@

Page 4 line 45: Define ``nodes" !

\textcolor{blue}{``Nodes'' is a standard word in the context of Bezier
  curves, defined in the caption of figure 2.  I have not used the
  word in any other sense in the manuscript and I think the text is
  clear as it stands.\\ \\}

@@

Figure 3, 5, 7: Perhaps fewer than 100 circles per Bezier curve would
be better.

\textcolor{blue}{Each coloured circle is linked to a Bezier point and
  this is set to 100 by default.  This is technically user-settable
  but to make it much fewer than 100 would seriously erode the quality
  of the strand plot.\\ \\}

Figure 4.  Knot $7_6$ with strands numbered... -- The main use here seems
to be to see very clearly the extent of every Bezier curve.
Over-/under-crossings seem to become relevant only for the final
rendering of the knot curve with the gaps included.

\textcolor{blue}{Good point, I have added a sentence to this effect to
  the caption of figure 4 which now ends with `` Crossings become
  relevant only for the final rendering of the knot curve with the
  gaps included''.  \\ \\}

@@

\textcolor{blue}{ An insightful observation.\\ \\ OLD TEXT In the
  broader context of optimisation in art, we observe that numerical
  optimisation techniques can produce aesthetically pleasing results,
  an observation that might find uptake by graphic artists.\\ \\ NEW
  TEXT In the broader context of optimisation in art, we observe that
  numerical optimisation techniques can produce aesthetically pleasing
  results, an observation that might find uptake by graphic artists.
  Indeed, one could argue that real art lies in the capturing of
  mathematical formulations of aesthetics.  \\ \\ }

Gallery:

Knot diagrams $8_3$ through $8_6$ in Figure 12 stand out with their helices
too tightly wound.  The two criteria on page 4, lines 54 and 55 do not
seem to be sufficiently enforced.  I suggest adjusting these figures
in the spirit of Figure 9b.  This may happen automatically with a
better ``badness function" that penalizes non-orthogonal crossing
angles more directly.

\textcolor{blue}{The referee is correct to say that changing the
  badness function would result in diagrams less like Figure 10a and
  more like 10b.  But I would take issue with the phrase ``too tightly
  wound''.  I have polled my colleagues who appear to be evenly split
  between preferring 10a and 10b.  The comparison appears to elicit
  strong opinions here at AUT.  Proponents of 10b point to the broader
  winding along the central helix, while those of 10a point to the
  more liquid and languid appearance, and in particular emphasise the
  absence of retrograde curvature at the bottom.  I have added a brief
  discussion of these issues to the caption of figure 12 [which is now
    figure 15 of the resubmission].  I would observe that the {\tt
    knotR} package makes such discussions possible.\\ \\}

### 2022-06-15 inst/jma_knot.tex
minor tweaks

Two key questions emerge.  Firstly, {\em What makes a knot plot
  attractive?}, and secondly {\em How best to create knot plots with
  the desired qualities?}  I address both questions, and attempt to
capture the aesthetics of knot diagrams using mathematical
formulations.  These formulations are of two types: hard constraints
of different forms of symmetry; and quantification of various
undesirable qualities of a knot diagram that may be minimized
numerically.  I present software and workflows that not only produce
pleasing knot diagrams, but also allow one to investigate the
connection between the aesthetics of knot diagrams and their
mathematics.

@@

One plausible technique for creating knot projections is to consider a
two-dimensional projection of a knot's embedding
$\mathbb{E}\in\mathbb{R}^3$.  However, this approach often results in
poor visual appearance: cusps or other displeasing effects can occur.

>>>>

Two key questions emerge.  Firstly, {\em What makes a knot plot
  attractive?}, and secondly, {\em How best to create knot plots with
  the desired qualities?}  I address both questions, and attempt to
capture the aesthetics of knot diagrams using mathematical
formulations.  These formulations are of two types: hard constraints
of different forms of symmetry; and quantification of various
undesirable qualities of a knot diagram that may be minimized
numerically.  I present software and workflows that not only produce
pleasing knot diagrams, but also allow one to investigate the
connection between the aesthetics of knot diagrams and their
mathematics.

@@

One plausible technique for creating knot projections is to consider a
two-dimensional projection of a knot's embedding
$\mathbb{E}\subset\mathbb{R}^3$.  However, this approach often results in
poor visual appearance: cusps or other displeasing effects can occur.
